---
title: "REU_UVR_Polyp"
author: "Lauren E Johnson"
date: "06/1/2022"
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### **UVR consequences on polyp reproduction**  
**Importing and tidying asexual reproduction data**  

Reproduction Results 
1. One-way ANOVA (parametric); Tukey Post-Hoc
2. Two-way ANOVA (parametric); Tukey Post-Hoc
3. Kruskall-Wallace (non-parametric); Pairwise Wilcox Test

Detachment Results
1. Chi-squaredtest of independence 


1. Importing data. 
```{r}
# data where all the controls are combined into one group 
df.combinedcontrol <- read.csv(here::here("data_raw/polypclonedata_combinedcontrol.csv"), stringsAsFactors = F)
```

2. Add column with final polyp numbers 
```{r message=FALSE, warning=FALSE}
df.combinedcontrol$FinalPolyp <- df.combinedcontrol$Day_27
```

3. Calculate daily and weekly budding rates
```{r}
# daily budding rate for experiment
df.combinedcontrol$DBR <- ((df.combinedcontrol$FinalPolyp - df.combinedcontrol$Day_00)/27)
# weekly budding rate for experiment
df.combinedcontrol$WBR <- (((df.combinedcontrol$FinalPolyp - df.combinedcontrol$Day_00)/27)*7)
```

4. Reorganize/Reformat the data frame 
```{r message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
```

```{r}
# long format 
df.combinedcontrol.gather <- df.combinedcontrol %>% gather(key = "Time", value = "Polyps", Day_00:Day_27)

# separate the days of experiment out and only keep the number of the day
df <- separate(data = df.combinedcontrol.gather, col = Time, into = c(NA,'Day'), sep = '_')

# reformat number of day to integer
df$Day <- as.integer(df$Day)

knitr::kable(head(df))

```


**Data analysis**  

We compared differences in weekly budding rate among the UVR treatment groups using a one-way ANOVA with treatment as the main effect.  

```{r}
wbraov <- aov(WBR~Treatment, data = df.combinedcontrol)

summary(wbraov)
```

A post-hoc Tukey test was used when significant differences were detected.   
```{r}
TukeyHSD(wbraov)
```
  
### **UVR consequences on polyp health**  
**Creating table of detachment data** 

Polyp detachment was classified binomially as either yes (polyps in dish detached at > 1 timepoints) or no (polyps in dish detached at ≤ 1 timepoints).  
```{r}
D <- as.table(rbind(c(11,7), c(11,1), c(4,8), c(0,12)))

dimnames(D) <- list(treatment = c("Control", "AA", "BB", "AB"), status = c("Attached", "Detached"))

knitr::kable(D)
```

**Data analysis**    

Chi-squared test of independence to test whether there were non-random differences in detachment probabilities across the treatment groups.  
```{r message=FALSE, warning=FALSE}
library(chisq.posthoc.test)

chisq.test(D, correct = F)

```

Post-Hoc Comparisons  
  
6 comparisons so alpha is 0.05/6 = 0.0083  
Comparison is bolded if p value is significant  

Control vs AA  
```{r warning=FALSE}
chisq.test(D[c(1,2),], correct = F)
```

Control vs BB  
```{r}
chisq.test(D[c(1,3),], correct = F)
```

**Control vs AB**  
```{r warning=FALSE}
chisq.test(D[c(1,4),], correct = F)
```

**AA vs BB**  
```{r warning=FALSE}
chisq.test(D[c(2,3),], correct = F)
```

**AA vs AB**  
```{r}
chisq.test(D[c(2,4),], correct = F)
```

BB vs AB  
```{r warning=FALSE}
chisq.test(D[c(3,4),], correct = F)
```

### **Figures**  
**Fig 1A.** Total number of polyps observed (mean +/- s.e.) under each UV treatment over the 27-day experiment.   
```{r, echo=FALSE}

library(ggridges)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)
library(ggpubr)


ggline(df, x = "Day", y = "Polyps", 
       combine = T,
       color = "Treatment",
       add = c("mean_se"),
       add.params = list(),
       ylab = expression("Number of polyps (mean ± s.e.)"),
       xlab = "Day of Experiment",
       shape = "Treatment",
       size = .3,
       font.label = list(size = 30, color = "black"),
       ggtheme = theme_classic()) + scale_color_grey()

```
  
**Fig 1B.** Weekly budding rates (polyps/week) for each UV treatment. Lowercase letters (a, b, c) indicate signficant differences (p < 0.5) between groups.   
```{r, echo=FALSE}
# using df.combined control otherwise will have 27 points for each rate
ggplot(data = df.combinedcontrol, aes(x = Treatment, y = WBR)) +
  #stat_boxplot(geom = "errorbar", width = 0.2, show.legend = F) +
  geom_boxplot(color = "black", size = .3, show.legend = F, outlier.colour = NA, position=position_dodge(width = 2)) +
  geom_jitter(alpha = 0.5, width = 0.05, size = .75, color = "black", show.legend = F) +
  stat_summary(fun=mean, geom="point", color = "black", shape = 4, show.legend = F) +
  ylab('Weekly Buddling Rate') +
  theme_classic() 



```



### **Data Analysis in Response to Reviewers**
1. Importing data. Added UVA and UVB column independently. '1' indicates presence of that wavelength and '0' indicates absence. 
```{r}
# data where all the controls are combined into one group 
df <- read.csv(here::here("data_raw/polypclonedata_combinedcontrol_v2.csv"), stringsAsFactors = F)
```

2. Add column with final polyp numbers 
```{r message=FALSE, warning=FALSE}
df$FinalPolyp <- df$Day_27
```

3. Calculate daily and weekly budding rates
```{r}
# daily budding rate for experiment
df$DBR <- ((df$FinalPolyp - df$Day_00)/27)
# weekly budding rate for experiment
df$WBR <- (((df$FinalPolyp - df$Day_00)/27)*7)
```

4. Reorganize/Reformat the data frame 
```{r message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
```

```{r}

# separate the days of experiment out and only keep the number of the day
df <- separate(data = df, col = Time, into = c(NA,'Day'), sep = '_')

# reformat number of day to integer
df$Day <- as.integer(df$Day)

# Make UVA and UVB factors
df$UVA <- as.factor(df$UVA)
df$UVB <- as.factor(df$UVB)

knitr::kable(head(df))

```

**Data analysis**  
```{r}
interaction <- aov(WBR ~ UVA * UVB, data = df)

plot(interaction)
```

Shapiro-Wilk Test of Normality. WBR departs significantly from parametric assumption of normality.
```{r}
plot(interaction, 2)

rstatix::shapiro_test(residuals(interaction))
```

Levene's of homogeneity of variances. WBR departs significantly from parametric assumption.
```{r}
library(rstatix)
plot(interaction, 1)

levene_test(WBR~UVA * UVB, data = df)
```

Because the data failed parametric assumption, we decided to run an non-paramentric Kruskal-Wallis Rank Sum Test. Given that there was a signficant effect, we ran Pairwise Wilcox test with a Bonferroni correction to correct for multiple comparisons.
```{r}
kruskal.test(WBR ~ Treatment, data = df)

pairwise.wilcox.test(df$WBR, df$Treatment, p.adjust.method = "bonferroni")
```


We compared the relative effects of UVA and UVB on weekly budding rate using a two-way ANOVA, that included an interaction between UVA and UVB. We also created a model that considered JarID since it was a grouping factor. 
```{r}
interaction <- aov(WBR~UVA * UVB, data = df)
two.way <-aov(WBR~UVA + UVB, data = df)
blocking <- aov(WBR ~ UVA * UVB + JarID, data = df)
```

We compared the AICs of the models to see which is of best fit. 
```{r}
library(AICcmodavg)

model.set <- list(two.way, interaction, blocking)
model.names <- c("two.way", "interaction", "blocking")

aictab(model.set, modnames = model.names)
```

```{r}
summary(interaction)
```

A post-hoc Tukey test was used when significant differences were detected.   
```{r}
TukeyHSD(interaction)
```

```{r}
summary(blocking)
```

